SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-10-14 maintex.tex
Next round of revisions

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 175 star-forming galaxies in close pairs with those of mass-matched isolated galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.3 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy.
\end{abstract}

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 $R$/\reff\ and 2.5 $R$/\reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

Where $NUV-r$ is the color from SDSS's k-corrected absolute magnitude and $\mathcal{M}_i$ is the i-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0 \reff\ aperture to remove any passive galaxies which may have passed the CMD cut \citep{Cid-Fernandes:2011}. We further use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with an AGN from our sample. We use the emission line ratios, log([O\textsc{III}]/H$\beta$) and log([N\textsc{II}]/H$\alpha$), extracted from a 1.3 kpc aperture along with the maximum starburst line from \citet{Kewley:2001} as the demarcation between star forming branch and the AGN branch.

@@

In our previous work, \citet{Fu:2018}, using the 2168 unique galaxies form DR14 \citep{Abolfathi:2018}, we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

@@

The spectra of the identified objects is extracted through a 1\arcsec\ circular aperture and fitted assuming the MaNGA target's redshift and then manually sorted into the following categories; ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra are well modeled by {\sc spfit} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming by the CMD diagram. These requirements reduce our sample down to 87 star forming galaxies (SFGs).

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. After restricting the sample to SFGs, we have another 133 MaNGA targets with paired galaxies outside of the IFU. This give use a total of 175 MaNGA targets in our pair sample.

@@

Setting the above requirements, most paired galaxies will find between 20$-$100 control galaxies. Since we want each paired galaxy to be treated in a similar manner, we will down-select the total number of acquired control galaxies down to 20 controls randomly. Control galaxies may be selected for multiple pairs by the pipeline; most of the control galaxies are used at least once and the average number of times a given control is reused is between 4$-$6 times. In the cases where a paired galaxy does not initially acquire 20 control galaxies we iteratively expand the stellar mass limit by 0.1 dex until at least 20 control galaxies are found. Only 5 of the paired galaxies required an extra iteration to find 20 controls.

While the first method is less sophisticated in comparison to the second method, we will see in Section \ref{sec:results} that the first method is sufficient to see the central enhancement to the star formation rate and the mass independence of the star formation enhancement.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Radial Profiles of Star Formation}\label{sec:analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/ssfr_comb.pdf}
\caption[]{The log(sSFR) as a function of galactocentric radius for control galaxies (\textbf{Left}) and galaxy pairs (\textbf{Middle}). The difference between the profiles of the paired galaxies and the control galaxies are shown in the \textbf{Right} panel. The dashed black profile represents the mean of the difference profiles. The colors of the profiles represent the mass range of the selected galaxies which is given in the legend along with the number of galaxies in that mass bin in parantheses. The highlighted region around the profiles represent the standard error of the mean of the data at the given radius interval. The parts of the profiles with dashed lines represent regions where less than 50\% of the galaxies in the bin are covered at that radii. }
\label{fig:ssfr_prof}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Specific Star Formation Rate}

@@

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction in parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination.

\begin{equation}
A_V = 6.78\ {\rm log} \left[(H\alpha/H\beta)/2.85 \right]
\end{equation}

We measure the SFR from the extinction corrected H$\alpha$ luminosity, L$_{{\rm H}\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

Since the stellar mass of a galaxy is not uniformly distributed within the galaxy, we normalize the SFR by the local stellar mass, $M_*$, giving us the specific star formation rate (sSFR). The local stellar masses used here is the stellar mass calculated by {\sc SPFIT}'s fit to the stellar continuum within each spaxel (spatial pixel).

@@

This will allow us to compare regions to high mass surface density to regions of low mass surface density.

We check the results of the radial profiles of this sSFR against the radial profiles of \ewha\ and find that they are in good agreement. \ewha\ is an observable while our measurement of the sSFR is dependent on both the H$\alpha$ luminosity and {\sc spfit}'s calculation of the stellar mass, so this check helps confirm the quality of our sSFR measurement.

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. First, the geometry of the galaxies needs to be defined. Specifically we will need position angle, inclination angle, and the effective radius of each of the MaNGA targets. We use the $r$-band elliptical Petrosian apertures from the NSA catalog and the $r$-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies.

@@

Once the geometry of the galaxies are set, we can build the radial profiles. The spaxels are binned into radius increments of 0.2 R/\reff\ from 0.0$-$2.6 R/\reff. Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We create one of these individual averaged radial profiles for every MaNGA galaxy.

@@

With individual log(sSFR) profiles built for each MaNGA galaxy, we now use two different methods to compare the profiles of paired galaxies against the profiles of control galaxies. The first method, in Section \ref{sec:mass-bin}, paired galaxies are compared to control galaxies within evenly spaced stellar mass-bins. In the second method, in Section \ref{sec:tailored}, paired galaxies are compared to a subset of 20 control galaxies which have a similar stellar mass and redshift.

@@

The radial extent of our profiles go out to 2.6 $R$/\reff; however,  67\% of the MaNGA sample is designed to cover only 1.5 $R$/\reff\ of the galaxy. Also, the MaNGA galaxies are not always centered on a spaxel center, so the 0$^{\rm th}$ spaxel may fall into the 0.2 $R$\reff\ bin. Because of these issues, the centers or the outskirts of the stacked profiles may only represent a small subset of the galaxies in that mass bin. To account for this, we ``trim" the stacked profiles where radii bins contain data for less than 50\% of the total set of galaxies. We represent the trimmed regions in our figures with dashed lines. In Figure \ref{fig:ssfr_prof}, this is shown in the centers of the lower mass galaxies, \logm\ $\le$ 10.0.

Once the stacked profiles are made, we take the difference between the stacked profiles of the paired galaxies and the control galaxies, pair - control, in log space (this means that the profiles really represents a ratio between the pairs and controls in linear space). This gives us difference profile, $\Delta$log(sSFR), which shows us where the sSFR is enhanced or suppressed (shown in the right hand panel of Figure \ref{fig:ssfr_prof}).

%The dashed regions of the profiles in Figure \ref{fig:ssfr_prof} represent regions which have been ``trimmed" due to a low percent of data coverage in the region. A radial bin is trimmed if less than 70\% of the galaxies in that bin have data. There are a few reasons why galaxy's profile may not have data points at a given radial bin. First, the MaNGA IFU may not cover the radius range; only $\sim$37\% of MaNGA galaxies are covered out to 2.5 $R$/\reff. This radial coverage limitation primarily affect the lower mass galaxies in the \logm\ $=$ 9.0$-$9.5. Second, foreground and background objects have been masked out which can create gaps in the data of an individual galaxies profile. This case does not significantly affect the stacked profiles here since the masked out regions will be spread out across the whole radius range. Third, retired regions within galaxies have been masked out with the \ewha\ $\ge$ 3\AA\ cut. It is well known that high mass galaxies tend to be passive (as seen in our CMD diagram in Figure \ref{fig:cmd}). As a consequence, the highest mass galaxies in our sample (\logm\ $=$ 11.0$-$11.5) have much of their profiles trimmed due to absence of star forming spaxels.

The profiles of the control galaxies are lower in their centers, which was previously seen in the MaNGA sample in \citet{Belfiore:2018}. The profiles of the paired galaxies flat across the whole galaxy accept for galaxies with masses above \logm\ $\ge$ 10.5 which show some central quenching. Because of this, it may be more accurate to describe the control galaxies as centrally suppressed or quenched in comparison to the paired galaxies which feature flat profiles of star formation. The difference in the pair and control profiles seems to indicate that the mechanism for quenching in galaxies is the depletion of gases in their centers. The gas inflows from the merger events can then refuel the centers of these quenched galaxies and allow new star formation to occur.

When taking the difference between the pairs and controls, the paired profiles are centrally enhanced by $\sim$0.35 $\pm$ 0.1 dex which then falls to zero around 1.2 R/\reff. In the outskirts of the pairs beyond 1.2 R/\reff, the pairs feature lightly suppressed sSFR of 0.0$-$0.1 dex. All of the mass ranges except for the highest mass range, \logm\ $=$ 11.0$-$11.5, trend closely to the median profile. The most massive galaxies in the range, \logm\ $=$ 11.0$-$11.5, the enhancement to the sSFR within 0.5 R/\reff\ is higher than the median profile at 0.5$-$0.6 dex. The sSFR enhancement in the sample seems to mass independent except for the most massive galaxies which feature higher levels of enhancement.

The higher rate of star formation within interacting galaxies will mean that the stellar mass in the interacting galaxies will grow at a faster rate compared to secular mass growth. The fractional mass growth rate is simply the difference between the sSFR profiles of the paired and control galaxies (pairs - controls) multiplied by at timescale. We chose a timescale of 1 Gyr as an order of magnitude estimate of the interaction event \citep{Boylan-Kolchin:2008}, though the actual timescale of the merger induced starburst is likely somewhat shorter ($\sim$0.4 Gyr; \citet{Feng:2019}).

Since the $\Delta$log(sSFR) is independent of galaxy mass, the fractional mass change will depend on the stellar mass of the paired galaxy. Low mass galaxies, \logm\ $=$ 9.0$-$9.5, experience a fractional mass change of $\Delta M_{\rm merger}$/$M_{\rm total}$ $=$ 0.06 (6\%) while massive galaxies, \logm\ $=$ 11.0$-$11.5, experience a fractional mass change of $\Delta M_{\rm merger}$/$M_{\rm total}$ $=$ 0.005 (0.5\%). This show that the merger induced mass growth is greatest in low mass galaxies. Further, the merger induced mass growth is restricted to the centers of the paired galaxies which will result in galaxies which have more prominent bulges in their centers.

This method has the benefit of a high S/N control profile since each mass bin will have hundreds of control galaxies while the median control profiles of the second method will only have 20 controls. On the other hand, this method only controls for a single galaxy parameter, stellar mass, though it is still relevant that the central enhancement of the sSFR in pairs is able to be observed. Further the profiles can only be studied as a function of stellar mass so we are unable to study the profiles as a function of projected separation or mass ratio.

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}

In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in Section \ref{sec:control}. We take the individual azimuthally averaged profiles of the selected control galaxies and take the median value of the profiles at each radius bin to produce the average profile of the control galaxies.

% Is this necessary
%The error of the profile is the standard error of the mean of the individual profiles. We also take into account the error associated with the random selection of controls using a bootstrapping method to quantify the variation in the values of the stacked profiles over 1000 iterations. We estimate that this random selection contributes an error of about 0.1 dex.

@@

The difference profiles at this stage have now been controlled for stellar mass and redshift. We can freely stack these difference profiles to study their dependence on other parameter like the mass ratio and projected separation. We split the individual difference profiles into five evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.5 and stack the difference profiles within each mass bin. This gives us five difference profiles covering five different stellar mass ranges shown in the left hand panel of Figure \ref{fig:ssfr_mass}. The errors of the stacked profiles are the standard error of the mean of the difference profiles within each bin.

@@

To understand why these high mass galaxies behave differently from the rest of the sample, we compare the distribution of the projected separation between the whole sample and the galaxies in the highest mass bin in Figure \ref{fig:sep_hist}. The whole sample has a relatively flat separation distribution while 70\% of the highest mass galaxies in the sample are within 20 kpc. As we will show in Section \ref{sec:sep}, the central sSFR enhancement is strongest at small projected separation which means that higher sSFR in the centers of the massive galaxies may be driven by their close projected separations.

The central enhancement to the sSFR falls to zero around 1.2 $R$/\reff\ and the outskirts of the galaxies are lightly suppressed by 0.1 dex. The paired galaxies in the mass range, \logm\ $=$ 9.0$-$9.5, feature enhanced sSFR in their disks of 0.1$-$0.2 dex. As we will show in the next section, Section \ref{sec:dm}, the offsets seen in the disks of paired galaxies are largely driven by whether the galaxy is the more massive or less massive companion in the pair.

@@

\subsection{Dependency on Mass Ratio}\label{sec:dm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by mass ratio. }
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In the top left of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins; the primary and secondary companion of a major merger, $|\Delta$log(M)$|$ $<$ 0.5, and the primary and secondary companion of a minor merger, $|\Delta$log(M)$|$ $>$ 0.5. Regardless of the mass ratio, the centers of the paired galaxies are centrally enhanced, though major mergers have stronger enhancements in their centers by $\sim$0.2 dex. Secondary companions also feature slightly higher levels of enhancement than primary companions at the same mass ratios by $\sim$0.1 dex.

The sSFR offsets in the outskirts of the galaxy pairs show a dependence on the mass ratio. Primary companions show a zero to positive enhancement, 0.0$-$ 0.4 dex, to the sSFR in their outskirts. Secondary companions feature a suppression to the sSFR in their outskirts of about 0.1$-$0.4 dex.

@@

From the mass ratio, we see two different effects. First, we see that the central enhancement to the sSFR is strongest for pairs with 1:1 mass ratios. Second, we see that secondary companions show steeper enhancement profiles with higher levels of sSFR enhancement in their centers and stronger sSFR suppression in their disks with respect to primary companions. It is interesting to note that the central enhancement to sSFR is dependent on both the mass ratio and the projected separation (which we will demonstrate in the next section) while the sSFR offsets in their outskirts primarily depends on whether the galaxy is the primary or secondary companion of the pair.

@@

We look at the level of sSFR enhancement as a function of the projected separation between the two galaxies in Figure \ref{fig:ssfr_sep}. The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation where close galaxy pairs show higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 0.0$-$12.5, are generally $\sim$0.05$-$0.1 dex above the median and the profiles of the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 37.5$-$50.0, are generally $\sim$0.05$-$0.1 dex below the median.

>>>>

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 175 star-forming galaxies in close pairs with those of mass-matched isolated galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies is dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences no to a positive sSFR enhancement while the less massive companion experiences a sSFR suppression in their outskirts.
\end{abstract}

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

Where $NUV-r$ is the color from SDSS's k-corrected absolute magnitude and $\mathcal{M}_{\rm i}$ is the i-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0 \reff\ aperture to ensure that all passive galaxies are removed from the sample \citep{Cid-Fernandes:2011}. We further use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with an AGN in their centers from our sample. We use the emission line ratios, log([O\textsc{III}]/H$\beta$) and log([N\textsc{II}]/H$\alpha$), extracted from a 1.3 kpc aperture along with the maximum starburst line from \citet{Kewley:2001} as the demarcation between star forming branch and the AGN branch.

@@

In our previous work, \citet{Fu:2018}, using the 2168 unique galaxies from DR14 \citep{Abolfathi:2018}, we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

@@

The spectra of the identified objects is extracted through a 1\arcsec\ circular aperture and fitted assuming the MaNGA target's redshift and then manually sorted into the following categories; ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra are well modeled by {\sc spfit} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within approximately $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in Section \ref{sec:data}. These requirements reduce our sample down to 87 star forming galaxies (SFGs).

@@

From the NSA catalog's 641,409 galaxies, we find 492 galaxies which are paired to MaNGA targets. After restricting the sample to SFGs, we have another 133 MaNGA targets with paired galaxies outside of the IFU. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. This give use a total of 175 MaNGA targets in our pair sample.

@@

Setting the above requirements, most paired galaxies will find between 20$-$100 control galaxies. Since we want each paired galaxy to be treated in a similar manner, we will down-select the total number of acquired control galaxies down to 20 controls randomly. A given control galaxy may be selected for multiple pairs by the pipeline; most of the control galaxies are used at least once and the average number of times a given control is reused is between 4$-$6 times. In the cases where a paired galaxy does not initially acquire 20 control galaxies we iteratively expand the stellar mass limit by 0.1 dex until at least 20 control galaxies are found. Only 5 of the paired galaxies required an extra iteration to find 20 controls.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Radial Profiles of Star Formation}\label{sec:analysis}

\subsection{Specific Star Formation Rate}

@@

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction is parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination.

We measure the SFR from the extinction corrected H$\alpha$ luminosity, L$_{{\rm H}\alpha}$.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

Since the stellar mass of a galaxy is not uniformly distributed within the galaxy, we normalize the SFR by the local stellar mass, $M_*$, giving us the specific star formation rate (sSFR). The local stellar masses used here is the stellar mass calculated by {\sc SPFIT}'s fit to the stellar continuum within each MaNGA spaxel (spatial pixel).

@@

This will allow us to compare regions of high mass surface density to regions of low mass surface density.

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. First, the geometry of the galaxies needs to be defined. Specifically we will need the position angle, the inclination angle, and the effective radius of each of the MaNGA targets. We use the $r$-band elliptical Petrosian apertures from the NSA catalog and the $r$-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies.

@@

Once the geometry of the galaxies are set, we can build the radial profiles. The spaxels are binned into radius increments of 0.2 \reff\ from 0.0$-$2.6 \reff. Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We create one of these individual averaged radial profiles for every MaNGA galaxy.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/ssfr_comb.pdf}
\caption[]{The log(sSFR) as a function of galactocentric radius for control galaxies (\textbf{Left}) and galaxy pairs (\textbf{Middle}). The difference between the profiles of the paired galaxies and the control galaxies are shown in the \textbf{Right} panel. The dashed black profile represents the mean of the difference profiles. The colors of the profiles represent the mass range of the selected galaxies which is given in the legend along with the number of galaxies in that mass bin in parantheses. The highlighted region around the profiles represent the standard error of the mean of the data at the given radius interval. The parts of the profiles with dashed lines represent regions where less than 50\% of the galaxies in the bin are covered at that radii. }
\label{fig:ssfr_prof}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

With the individual log(sSFR) profiles built for each MaNGA galaxy, we now use two different methods to compare the profiles of paired galaxies against the profiles of control galaxies. The first method, in Section \ref{sec:mass-bin}, paired galaxies are compared to control galaxies within evenly spaced stellar mass-bins. In the second method, in Section \ref{sec:tailored}, paired galaxies are compared to a subset of 20 control galaxies which are matched in stellar mass and redshift.

@@

The radial extent of our profiles go out to 2.6 \reff; however,  67\% of the MaNGA sample is designed to cover only 1.5 \reff\ of the galaxy. Also, the MaNGA galaxies are not always centered on a spaxel center, so the 0$^{\rm th}$ spaxel may fall outside of the innermost radii bin. Because of these issues, the centers or the outskirts of the stacked profiles may only represent a small subset of the galaxies in that mass bin. To account for this, we ``trim" the stacked profiles where a radius bin contains data for less than 50\% of the total set of galaxies. We represent the trimmed regions in our figures with dashed lines. In Figure \ref{fig:ssfr_prof}, galaxies below \logm\ $\le$ 10.0 have their innermost radii bin trimmed due to insufficient coverage.

Once the stacked profiles are made, we take the difference between the stacked profiles of the paired galaxies and the control galaxies, pair - control, in log space (this means that the difference profiles really represent a ratio between the pairs and controls in linear space). This gives us difference profile, $\Delta$log(sSFR), which shows us where the sSFR is enhanced or suppressed (shown in the right hand panel of Figure \ref{fig:ssfr_prof}).

When taking the difference between the pairs and controls, the paired profiles are centrally enhanced by $\sim$0.35 $\pm$ 0.1 dex which then falls to zero around 1.2 \reff. In the outskirts of the pairs beyond 1.2 \reff, the pairs feature lightly suppressed sSFR of 0.0$-$0.1 dex. All of the mass ranges except for the highest mass range, \logm\ $=$ 11.0$-$11.5, trend closely to the median profile. For the most massive galaxies, \logm\ $=$ 11.0$-$11.5, the enhancement to the sSFR within 0.5 \reff\ is higher than the median profile at 0.5$-$0.6 dex. The sSFR enhancement in the sample seems to be mass independent except for the most massive galaxies which feature higher levels of sSFR enhancement.

The higher rate of star formation within interacting galaxies will mean that the stellar mass in the interacting galaxies will grow at a faster rate compared to secular mass growth. The fractional mass growth rate is simply the difference between the sSFR profiles of the paired and control galaxies in linear space (pairs - controls) multiplied by a given timescale. We chose a timescale of 0.4 Gyr which is estimated to be the average timescale for the merger induced starburst \citep{Feng:2019}.

Since the $\Delta$log(sSFR) is independent of galaxy mass, the fractional mass change will depend on the stellar mass of the paired galaxy. That is, since the amount of new stellar mass from merger induced interactions is roughly constant across galaxies of different stellar masses, lower mass galaxies will experience a greater change to their total stellar mass. Low mass galaxies, \logm\ $=$ 9.0$-$9.5, experience a fractional mass change of $\Delta M_{\rm merger}$/$M_{\rm total}$ $=$ 0.03 (3\%) while massive galaxies, \logm\ $=$ 11.0$-$11.5, experience a fractional mass change of $\Delta M_{\rm merger}$/$M_{\rm total}$ $=$ 0.001 (0.1\%). This shows that the merger induced mass growth is greatest in low mass galaxies. Further, the merger induced mass growth is restricted to the centers of the paired galaxies which will result in galaxies which have more prominent bulges in their centers.

This mass-binning method has the benefit of a high S/N control profile since each mass bin will have hundreds of control galaxies while the median control profiles of the second method will only have 20 controls. On the other hand, this method only controls for a single galaxy parameter, stellar mass, though it is still relevant that the central enhancement of the sSFR in pairs is able to be observed. The profiles can also only be studied as a function of stellar mass so we are unable to study the profiles as a function of projected separation or mass ratio.

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by mass ratio. }
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in Section \ref{sec:control}. We take the individual azimuthally averaged profiles of the selected control galaxies and take the median value of the profiles at each radius bin to produce the median profile of the control galaxies.

@@

The difference profiles at this stage have now been controlled for both stellar mass and redshift. We can freely stack these difference profiles to study their dependence on other parameter like the mass ratio and projected separation.

First, we split the profiles by stellar mass to see how this method compares to the previous mass-binning method. We split the individual difference profiles into five evenly spaced stellar mass bins between \logm\ $=$ 9.0$-$11.5 and stack the difference profiles within each mass bin. This gives us five difference profiles covering five different stellar mass ranges shown in the left hand panel of Figure \ref{fig:ssfr_mass}. The errors of the stacked profiles are the standard error of the mean of the difference profiles within each bin.

@@

The central enhancement to the sSFR falls to zero around 1.2 \reff\ and the outskirts of the galaxies are lightly suppressed by 0.1 dex. The paired galaxies in the mass range, \logm\ $=$ 9.0$-$9.5, feature enhanced sSFR in their disks of 0.1$-$0.2 dex. As we will show in the next section, Section \ref{sec:dm}, the offsets seen in the disks of paired galaxies are largely driven by whether the galaxy is the more massive or less massive companion in the pair.

To understand why these high mass galaxies behave differently from the rest of the sample, we compare the distribution of the projected separation for the whole sample and for galaxies in the highest mass bin in Figure \ref{fig:sep_hist}. The whole sample has a v-shaped separation distribution. This is the result of the combination of the two different samples, the inside-IFU and outside-IFU samples. The outside-IFU sample should find more pairs at wide projected separations due to the larger volume to find companions in. The separation distribution of the inside-IFU sample is limited by MaNGA's IFU sizes and the sample's redshift distribution; the maximum possible projected separation for a MaNGA target with $z$ $=$ 0.15 and a 127 spaxel IFU is about 40 kpc. The highest mass galaxies in the sample are at closer projected separations in comparison to the rest of the sample with 70\% of the highest mass galaxies being within 20 kpc. As we will show in Section \ref{sec:sep}, the central sSFR enhancement is strongest at small projected separation which means that higher sSFR in the centers of the massive galaxies may be driven by their close projected separations.

@@

\subsection{Dependency on Mass Ratio}\label{sec:dm}

@@

In the top left of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins; the primary and secondary companion of a major merger, $|\Delta$log(M)$|$ $<$ 0.5, and the primary and secondary companion of a minor merger, $|\Delta$log(M)$|$ $>$ 0.5. Regardless of the mass ratio, the centers of the paired galaxies are centrally enhanced, though major mergers have stronger enhancements in their centers by $\sim$0.2 dex over minor mergers. Secondary companions also feature slightly higher levels of enhancement in their centers than primary companions at the same mass ratios by $\sim$0.1 dex.

The sSFR offsets in the outskirts of the galaxy pairs show a dependence on the mass ratio. Primary companions show a zero to positive enhancement of 0.0$-$0.4 dex to the sSFR in their outskirts. Secondary companions feature a suppression to the sSFR in their outskirts of about 0.1$-$0.4 dex.

@@

From the mass ratio, we see two different effects. First, we see that the central enhancement to the sSFR is strongest for pairs with 1:1 mass ratios. Second, we see that secondary companions show steeper enhancement profiles with higher levels of sSFR enhancement in their centers and stronger levels of sSFR suppression in their disks with respect to primary companions. It is interesting to note that the central enhancement to sSFR is dependent on both the mass ratio and the projected separation (which we will demonstrate in the next section) while the sSFR offsets in their outskirts primarily depends on whether the galaxy is the primary or secondary companion of the pair.

@@

We look at the level of sSFR enhancement as a function of the projected separation between the two galaxies in Figure \ref{fig:ssfr_sep}. The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation with close galaxy pairs showing higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 0.0$-$12.5, are generally $\sim$0.05$-$0.1 dex above the median and the profiles of the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 37.5$-$50.0, are generally $\sim$0.05$-$0.1 dex below the median.
